---
title: "ESM 204 Homework 2"
author: "Grace Bianchi, Claire Meuter, Chase Tarr"
date: "4/19/2022"
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE, warning = FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(here)
library(tidyverse)
library(janitor)
```

```{r, message=FALSE, warning = FALSE}
# read in data
water_districts <- read_csv(here("Water_Districts.csv")) %>% 
  clean_names() %>% 
  mutate(kern_e = 150 - reduction) %>% 
  mutate(mojave_e = 140 - reduction) %>% 
  mutate(antelope_e = 220 - reduction) %>% 
  mutate(ventura_e = 245 - reduction)

# linear regression with zero intercept
kern_lm <-lm(mc_kern ~ 0 + reduction, data = water_districts) # coefficient = 2.286
mojave_lm <-lm(mc_mojave ~ 0 + reduction, data = water_districts) # coefficient = 3.804
antelope_lm <-lm(mc_antelope ~ 0 + reduction, data = water_districts) # coefficient = 2.858
ventura_lm <-lm(mc_ventura ~ 0 + reduction, data = water_districts) # coefficients = 1.776
```

### 1. For each irrigation district, plot the marginal abatement cost data (from the data provided) and estimate a linear regression model with zero intercept. These estimated linear regressions will be your estimates of the marginal abatement cost curve (one for each irrigation district), which you will use for the remainder of this analysis.

```{r, fig.align='center'}
# marginal abatement cost problem 1 
ggplot(data = water_districts, aes(x = reduction)) +
  # add kern points and linear regression
  geom_point(aes(y = mc_kern, color = "Kern")) +
  geom_abline(intercept = 0, slope = kern_lm$coefficient[1],) +
  # add mojave data and linear regression
  geom_point(aes(y = mc_mojave, color = "Mojave")) +
  geom_abline(intercept = 0, slope = mojave_lm$coefficient[1]) +
  # 
  geom_point(aes(y = mc_antelope, color = "Antelope")) +
  geom_abline(intercept = 0, slope = antelope_lm$coefficient[1]) +
  #
  geom_point(aes(y = mc_ventura, color = "Ventura")) +
  geom_abline(intercept = 0, slope = ventura_lm$coefficient[1]) +
  labs(x = "Quantity of Abatement (acre feet)",
       y = "Marginal Cost ($/acre feet)",
       title = "Marginal Abatement Cost of Irrigation Districts") +
  scale_color_manual(name='Water District',
                     breaks=c('Kern', 'Mojave', 'Antelope', 'Ventura'),
                     values=c('Kern'='cornflowerblue', 'Mojave'='chocolate1', 'Antelope'='darkolivegreen3', 'Ventura' = 'palevioletred')) +
  theme_bw() +
  theme(legend.position = c(0.1, 0.75))


```

The estimated linear regression for each district with a zero intercept,
where A is the reduction in water use in acre feet--

-   Kern: Marginal Cost of abatement = 2.286A

-   Mojave: Marginal Cost of Abatement = 3.804A

-   Antelope: Marginal Cost of Abatement = 2.858A

-   Ventura: Marginal Cost of Abatement = 1.776A

### 2. Using your estimated marginal abatement cost curves, derive each district's demand curve for water. In other words, how much would each district be willing to pay for the right to use the first AF of water, second AF, etc.? Plot each district's demand curve on a single graph. Which sector is willing to pay the most for the first AF of water?

```{r, message=FALSE, warning = FALSE}
# find intercepts
kern_slope <-kern_lm$coefficient[1]
k_int <- kern_slope * (150)

mojave_slope <-mojave_lm$coefficient[1]
m_int <- mojave_slope * (140)

antelope_slope <-antelope_lm$coefficient[1]
a_int <- antelope_slope * (220)

ventura_slope <-ventura_lm$coefficient[1]
v_int <- ventura_slope * (245)

# create functions for each demand equation slope(e0-e)
k_demand <- function(x){return(k_int-kern_slope*x)}
m_demand <- function(x){return(m_int-mojave_slope*x)}
a_demand <- function(x){return(a_int-antelope_slope*x)}
v_demand <- function(x){return(v_int-ventura_slope*x)}
```

```{r, message=FALSE, warning = FALSE, fig.align='center'}
# demand curve functions
ggplot(data = water_districts, aes(x=x)) +
  stat_function(fun = k_demand, aes(x = kern_e, color = "Kern")) +
  stat_function(fun = m_demand, aes(x = mojave_e, color = 'Mojave')) +
  stat_function(fun = a_demand, aes(x = antelope_e, color = "Antelope")) +
  stat_function(fun = v_demand, aes(x = ventura_e, color = 'Ventura')) +
  theme_bw() +
  labs(x = "Quantity of water (acre feet)",
       y = "Marginal Benefit ($/acre feet)",
       title = "Water Demand in Irrigation Districts",
       color = "Water District") +
  ylim(0, 650) + xlim(0, 250) +
    scale_color_manual(name='Water District',
                     breaks=c('Kern', 'Mojave', 'Antelope', 'Ventura'),
                     values=c('Kern'='cornflowerblue', 'Mojave'='chocolate1', 'Antelope'='darkolivegreen3', 'Ventura' = 'palevioletred')) +
  theme(legend.position = c(0.8, 0.7))

```

**Antelope district is willing to pay the most (\$650/AF) for the first
acre feet of water due to it having the highest y-intercept.**

### 3. Here you will analyze three alternative policies for reducing water use among these irrigation districts. In total, these irrigation districts will need to reduce water consumption from the current 755 AF down to 500 AF. For each intervention listed below, perform the following calculations: (1) calculate the stringency of the policy (defined below for each intervention) required to achieve the 500 AF target, (2) calculate the total cost of meeting the target using that approach, (3) calculate the cost to each district, and (4) calculate the tax revenue generated.


#### **Cap without trade scenario**

1.  In order to reduce the current consumption to 500 AF, each district
    must reduce their consumption by x:

    \- 755-500 = 255

    \- x = $\frac{255}{755}$ = 0.3377

    **Each district must reduce their water use by a magnitude of the fraction $\frac{255}{755}$ or 0.3377**

<!-- -->

2.  The total cost is calculated using the abatement curve by
    multiplying each districts' baseline by the stringency (0.3377) then
    calculating the marginal cost of abatement. The cost for each
    district calculated using the area under the abatement curve (area
    of triangle) and summed to give the **total cost of abatement =
    \$21,780.69**

-   **Kern cost of abatement = \$2,933.15**

    -   Q Abatement: 150 \* 0.3377 = 50.66 AF

    -   Marginal Cost of Abatement = 2.286A = 50.66 \* 2.286 =
        \$115.80/AF

    -   cost of Abatement (area of triangle): 115.80 \* 50.66 \* 0.5 =
        \$2,933.15

-   **Mojave cost of abatement: \$4,251.55**

    -   Q abatement: 140 \* 0.3377 = 47.28 AF

    -   Marginal Cost of Abatement = 3.804A = 47.28 \* 3.804 =
        \$179.85/AF

    -   cost of Abatement: 179.85 \* 47.28 \* 0.5 = \$4,251.55

-   **Antelope cost of abatement: \$7,887.08**

    -   Q abatement: 220 \* 0.3377 = 74.29 AF

    -   Marginal Cost of Abatement = 2.858A = 74.29 \* 2.858 =
        \$212.33/AF

    -   cost of Abatement: 74.29 \* 212.33 \* 0.5 = \$7,887.08

-   **Ventura cost of abatement: \$6,078.91**

    -   Q abatement: 245 \* 0.3377 = 82.74 AF

    -   Marginal Cost of Abatement = 1.776A = 1.776 \* 82.74 =
        \$146.94/AF

    -   cost of Abatement: 82.74 \* 146.94 \* 0.5 = \$6,078.91

4. There is no government revenue because there is just a cap on the water, not a tax.

#### Tax on water use

**Stringency:**
The horizontal sum is calculated by solving for Q: 

Kern MC = 2.286Q  --> Kern Q = MC / 2.286

-   Mojave Q = MC / 3.804

-   Antelope Q = MC / 2.858

-   Ventura Q = MC /1.776

The aggregate is Q= (1/2.286 + 1/3.804 + 1/2.858 + 1/1.776) * MC-> Q=1.613MC
Using the abatement curve, we can plug in 255 for Q. 255 = 1.613 * MC --> MC = 158.063, therefore **the stringency or the magnitude of the tax is $158.03/AF.**

The total cost = cost of abatement + the cost of the tax on the water used. 
district tax cost of water used: (Baseline - abatement) * tax
The amount of water abated is calculated by using the cost of the tax and solving for Q for each district's slope.  **The total cost is $97,527.74**
-   Kern: 158.063 = 2.286Q --> Q = 69.144 AF-> 6390.171 + 5464.554  **cost to Kern = $17,596.91**

-   Mojave: 158.063 = 3.804Q --> 41.552 AF, **cost to Mojave = $18,057.32**

-   Antelope: 158.063 = 2.858Q --> 55.305 AF, **cost to Antelope = $30,181.80**

-   Ventura: 158.063 = 1.776Q --> 88.999 AF, **cost to Ventura = $31,691.71**

The tax revenue is 500 * 158.063
**Tax revenue: $79,031.50**

#### Cap and trade
The stringency will be the same as the first scenario; each district must reduce their water use by a magnitude of the fraction $\frac{255}{755}$ = **0.3377** 

The marginal cost of abatement is 158.063 using the last unit abated (255 AF) and the optimal quantity is equal to that found in the tax scenario. The district cost is the difference between that quanity and the quantity found in the cap no trade scenario.  **The total cost is $38,742.98**

Kern: 158.063 = 2.286Q -> Q = 69.144 - 50.66 AF = selling 18.484
 **Kern's cost: (69.144 * 158.063 * 0.5) - (18.484 * 158.063) = $2,542.92**
Mojave: 158.063 = 3.804Q --> 41.552 AF - 47.28 AF = buying 5.728
  **Mojave's cost: (47.28 * 158.063) + (5.728 * 158.063) = $8,378.60**
Antelope: 158.063 = 2.858Q --> 55.305 AF - 74.29 AF = buying 18.985
  **Antelope's cost: (74.29 * 158.063) + (18.985 * 158.063) = $14,743.33**
Ventura: 158.063 = 1.776Q --> 88.999 AF - 82.74 AF = selling 6.259
  **Ventura's cost: (88.999 * 158.063) - (6.259 * 158.063) = $13,078.13**
  
###4 A severe drought hits California, and it is necessary to reduce water use from 500 AF down to 300 AF. Your job is to estimate the cost of the drought (i.e., the cost of reducing water use from 500 AF to 300 AF) to each irrigation district under each policy. Considering the entire jump from status quo (755 AF) to the drought (300 AF), which policy is preferred by each irrigation district? How does your answer depend on how rights are allocated to each district (in policies (a) and (c))2?

Policy a) Cap Without Trade Drought Cost [500 AF down to 300 AF]

Each district must reduce their water use by a magnitude of the fraction $\frac{300}{500}$ or 0.4.

-   **Kern cost of abatement = $1,804.92**
  New 500 AF Baseline Q: 99.345
      (99.345 * 0.4) = 39.738 AF
      MCA: (39.738 * 2.286) = $90.84
      TCA: (90.84 * 39.738) / 2 = $1,804.92
-   **Mojave Cost of Abatement = $2,616.35**
  New 500 AF Baseline Q: 92.722
      (92.722 * 0.4) = 37.0888 AF
      MCA: (37.0888 * 3.804) = $141.09
      TCA: (141.09 * 37.0888) / 2 = $2,616.35
-   **Antelope Cost of Abatement = $4,854.08**
  New 500 AF Baseline Q: 145.706
      (145.706* 0.4) = 58.2824 AF
      MCA: (58.2824 * 2.858) = $166.57
      TCA: (166.57 * 58.2824) / 2 = $4,854.08
-   **Ventura Cost of Abatement = $3,740.73**
  New 500 AF Baseline Q: 162.2635
      (162.2635 * 0.4) = 64.904 AF
      MCA: (64.904 * 1.776) = $115.27
      TCA: (115.27 * 64.904) / 2 = $3,740.73
    
    Total Cost is: $13,016.08

Cap Without Trade Drought Cost [755 AF down to 300]

Each district must reduce their water use by a magnitude of the fraction $\frac{300}{755}$ or 0.6026.

-   **Kern cost of abatement = $9,338.71**
  Original 755 AF Baseline Q: 150 AF
      (150 * 0.6026) = 90.39 AF
      MCA: (90.39 * 2.286) = $206.63
      TCA: (206.63 * 90.39) / 2 = $9,338.71
-   **Mojave Cost of Abatement = $13,537.08**
  Original 755 AF Baseline Q: 140 AF
      (140 * 0.6026) = 84.364 AF
      MCA: (84.364 * 3.804) = $320.92
      TCA: (320.92 * 84.364 / 2 = $13,537.08
-   **Antelope Cost of Abatement = $25,115.15**
  Original 755 AF Baseline Q: 220 AF
      (220 * 0.6026) = 132.572 AF
      MCA: (132.572 * 2.858) = $378.89
      TCA: (378.89 * 132.572) / 2 = $25,115.15
-   **Ventura Cost of Abatement = $19,355.46**
  Original 755 AF Baseline Q: 245 AF
      (245 * 0.6026) = 147.637 AF
      MCA: (147.637 * 1.776) = $262.20
      TCA: (262.20 * 147.637) / 2 = $19,355.46
      
      Total Cost is: $67,346.40

Policy b) Taxing Water Use

[500 AF to 300 AF]
The aggregate is Q= (1/2.286 + 1/3.804 + 1/2.858 + 1/1.776) * MC-> Q=1.613MC
Using the abatement curve, we can plug in 200 for Q. 200 = 1.613 * MC --> MC = 123.993, therefore **the stringency or the magnitude of the tax is $123.99/AF.**

  **Tax Revenue -> 200 * 123.99 = $24,798.00**

[755 AF to 300 AF]
The aggregate is Q= (1/2.286 + 1/3.804 + 1/2.858 + 1/1.776) * MC-> Q=1.613MC
Using the abatement curve, we can plug in 455 for Q. 455 = 1613* MC --> MC = 282.083, therefore **the stringency or the magnitude of the tax is $282.08/AF.**

  **Tax Revenue -> 455 * 282.08 = $128,346.40**

Policy c) Cap and Trade

[500 AF to 300 AF]

The Stringency will be the same as above for this reduction. Each district must reduce their water use by a magnitude of the fraction $\frac{300}{500}$ or 0.4.

The marginal cost of abatement is $123.99 using the last unit abated (200 AF) and the optimal quantity is equal to that found in the tax scenario. The district cost is the difference between that quantity and the quantity found in the cap no trade scenario.
 **The Total Cost is $23,841.66**

Kern: 123.99 = 2.286Q -> Q = 54.239 - 39.738 AF = selling 14.501
 **Kern's cost: (54.239 * 123.99 * 0.5) - (14.501 * 123.99) = $1,564.57**
Mojave: 123.99 = 3.804Q --> 32.595 AF - 37.0888 AF = buying 4.4938
  **Mojave's cost: (37.0888 * 123.99) + (4.4938 * 123.99) = $5,155.83**
Antelope: 123.99 = 2.858Q --> 43.383 AF - 58.2824 AF = buying 14.8994
  **Antelope's cost: (58.2824 * 123.99) + (14.8994 * 123.99) = $9,073.81**
Ventura: 123.99 = 1.776Q --> 69.814 AF - 64.904 AF = selling 4.910
  **Ventura's cost: (69.814 * 123.99) - (4.910 * 123.99) = $8,047.45**

[755 AF to 300 AF]

The Stringency will be the same as above for this reduction. Each district must reduce their water use by a magnitude of the fraction $\frac{300}{755}$ or 0.6026.

The marginal cost of abatement is $282.08 using the last unit abated (455 AF) and the optimal quantity is equal to that found in the tax scenario. The district cost is the difference between that quantity and the quantity found in the cap no trade scenario.
 **The Total Cost is $123,373.52**

Kern: 282.083 = 2.286Q -> Q = 123.396 - 90.39 AF = selling 33.006
 **Kern's cost: (123.396 * 282.083* 0.5) - (33.006 * 282.083) = $8,093.53**
Mojave: 282.083 = 3.804Q --> 74.154 AF - 84.364 AF = buying 10.210
  **Mojave's cost: (84.364 * 282.083) + (10.210 * 282.083) = $26,677.72**
Antelope: 282.083 = 2.858Q --> 98.681 AF - 132.572 AF = buying 33.891
  **Antelope's cost: (132.572 * 282.083) + (33.891 * 282.083) = $46,956.38**
Ventura: 282.083 = 1.776Q --> 158.831 AF - 147.637 AF = selling 11.194
  **Ventura's cost: (158.831 * 282.083) - (11.194 * 282.083) = $41,645.89**


For Kern Irrigation District, if going from 500 AF to 300 AF across the market, Cap and Trade benefits them the most as they are going to be able to sell water usage rights to other districts and profit **$1,564.57**. If going from the original 755 AF baseline to 300 AF across the market, they would choose Cap and Trade once again as they will be able to sell units of water profiting **$8,093.53**.

For Mojave Irrigation District, if going from 500 AF to 300 AF across the market, they would rather have Cap without Trade as their cost would be **$2,616.35 instead of the Cap and trade cost of $5,155.83**. If going from the original 755 AF baseline to 300 AF across the market, they would choose Cap without Trade once again as their cost would be **$13,537.08 compared to having to buy $26,677.72** worth of water units in the Cap and Trade market from another irrigation district.

For Antelope Irrigation District, they are going to choose similar to Mojave if going from 500 AF to 300 AF across the market, they would rather have Cap without Trade as their cost would be **$4,854.08 instead of the Cap and trade cost of $9,073.81**. If going from the original 755 AF baseline to 300 AF across the market, they would choose Cap without Trade once again as their cost would be **$19,355.46 compared to having to buy $46,956.38** worth of water units in the Cap and Trade market from another irrigation district.

For Ventura Irrigation District, they are going to choose similar to Kern if going from 500 AF to 300 AF across the market, Cap and Trade benefits them the most as they are going to be able to sell water usage rights to other districts and have costs near **$8,047.45**. If going from the original 755 AF baseline to 300 AF across the market, they would choose Cap and Trade once again as they will be able to keep costs near due to selling units of water resulting in a cost of **$41,645.89**.





